```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part0: Overview
```{r Import data set, warning = FALSE}
library(readr)
Movies.original <- read_csv("tmdb_5000_movies.csv",show_col_types = FALSE)
# head(Movies.original) 

Movies.related <- subset(Movies.original, select = -c(homepage, id, overview, tagline, original_title, production_countries, keywords, spoken_languages, status)) #Delete irrelevant columns

DuplicatedIndex <- duplicated(Movies.related) #Check for repeated information
num_Duplicate <- which(DuplicatedIndex) #No duplicated movie

Movies.organized <- data.frame(na.omit(Movies.related)) #Omit rows with NAs
```


# Part1: Data Visualization

### 1.1 Categorical Features

##### 1.1.1 Function of Counting Categories
```{r Function of Counting Categories}
library("stringr")
Count.Category <- function(category){
  count_name = c()
  count_nums = c()
  for (i in 1:nrow(Movies.organized)){
    category_split = unlist(str_split(category[i], "\"")) # variable that stores the category names
    for (k in 1:length(category_split)){
      if (category_split[k]==": "){ 
        category_name = category_split[k+1]
        if (category_name %in% count_name){# if the category name already exists, add one
          count_nums[which(count_name == category_name)] = count_nums[which(count_name == category_name)]+1
        }
        else{ # else, add the category name to the list of category
          count_name = c(count_name,category_name)
          count_nums[which(count_name == category_name)] = 1
        }
      }
    }
  }
  names(count_nums) <- count_name # assign corresponding category names to the numbers
  return(count_nums)
}
```

##### 1.1.2 Distribution of Release Year
```{r Histogram of Release Year, message = FALSE}
release_year <- as.numeric(substring(Movies.organized$release_date, 1,4))
release_year.sort <- sort(table(release_year))
release_year.result <- summary(Movies.organized$release_date)
hist(release_year, col = "lightblue", xlim = range(1916, 2017), ylim =range(0, 500), breaks = 40, ylab = 'number of movies', main ='Histogram of Release Year')
# consider movies from 2000 to the most current data
Movies.organized$release_year <- release_year
Movies.organized <- Movies.organized[which(Movies.organized$release_year==2009),]
```

##### 1.1.3 Distribution of Split Genres
```{r Bar Plot of Split Genres}
genre_count_nums <- Count.Category(Movies.organized$genres)
barplot(sort(genre_count_nums,decreasing=TRUE),las = 2,col="lightblue", main="Bar Plot of Genres",cex.names=0.7)
```

##### 1.1.4 Distribution of Top 5 Movie Production Companies
```{r Pie Chart of Top 5 Companies}
company_count_nums <- Count.Category(Movies.organized$production_companies)
company_5 <- head(sort(company_count_nums,decreasing = T),5) # choose the top 5 companies that produce the largest number of movies
pie(company_5,main="Pie Plot of Top 5 Movie Production Companies",cex=0.7)
```

##### 1.1.5 Distribution of Languages
```{r Distribution of Languages}
library(ggplot2)
ggplot(Movies.organized,aes(x=original_language, y=popularity, color=original_language))+
  ggtitle("Scatter Plot of Languages")+
  geom_point()+
  theme(legend.title=element_text(size = 9),legend.text=element_text(size = 9))
```

### 1.2 Correlation

##### 1.2.1 Scatter Plot Matrix
```{r Scatter Plot Matrix}
pairs(~popularity+budget+revenue+runtime+vote_average+vote_count, data = Movies.organized, main = "Scatter Plot Matrix", panel = panel.smooth)
```

##### 1.2.2 Correlation Table
```{r Correlation Table}
res <- cor(data.frame(subset(Movies.organized, select = c(budget, revenue, runtime, vote_average, vote_count))))
```


#Part2: Data Processing

### 2.1 Processing Missing Values

##### 2.1.1 Numeric Variables
```{r Function Check Zeros}
check_zeros <- function(check_col){
  num_zeros <- length(which(check_col == 0))
  percentage_zeros <- num_zeros/nrow(check_col)
  return(c(num_zeros,percentage_zeros))
}
```

```{r Select numeric columns and apply function}
num_cols <- which(unlist(lapply(Movies.organized, is.numeric))) 
zeros_nums <- list()
zeros_ratio <- list()
for(i in 1:length(num_cols)){
  check_col <- Movies.organized[,num_cols[i]]
  zeros_nums[i] <- check_zeros(check_col)[1]
  zeros_ratio[i] <- check_zeros(check_col)[2]
}
names(zeros_nums) = names(num_cols)
zeroResult <- cbind(zeros_nums,zeros_ratio) 
```

```{r Removing rows with zero value}
Movies.organized <- Movies.organized[which(!(Movies.organized$budget==0)),]
Movies.organized <- Movies.organized[which(!(Movies.organized$revenue==0)),]
Movies.organized <- Movies.organized[which(!(Movies.organized$runtime==0)),]
Movies.organized <- Movies.organized[which(!(Movies.organized$vote_average==0)),]
Movies.organized <- Movies.organized[which(!(Movies.organized$vote_count==0)),]
Movies.organized.result <- summary(Movies.organized)
```

##### 2.1.2 Categorical Variables
```{r Removing NAs in categorical variables}
Movies.organized <- Movies.organized[which(!(Movies.organized$genres=='[]')),]
Movies.organized <- Movies.organized[which(!(Movies.organized$production_companies=='[]')),]
```

### 2.2 Dummy Variables
```{r Function of split}
split.category = list()
spe_category <- function(category){
  for(i in 1:nrow(Movies.organized)){
    category_split = unlist(str_split(category[i], "\""))
    spe_list = c()
    for(k in 1:length(category_split)){
      if(category_split[k] == ": "){
        spe_list = c(spe_list,category_split[k+1])
      }
    }
    split.category[[i]] = spe_list
  }
  return(split.category)
}
```

##### 2.2.1 Genre
```{r Data set for regression}
Movies <- data.frame(subset(Movies.organized, select = c(popularity,budget, revenue, runtime,vote_average, vote_count))) # create new data set with only numeric variables
```

```{r Split Genres}
genre_count_nums <- Count.Category(Movies.organized$genres)
genre_count.sort <- sort(genre_count_nums,decreasing=TRUE)
genreNames.factor <- as.factor(names(genre_count_nums))

split.genres <- spe_category(Movies.organized$genres)
Movies.organized$genres <- split.genres # replace genres with a more readable form
```

```{r Create dummy variables for genres}
Movies$dDrama <- 0 #(Dummy Drama)
for (i in 1:length(Movies.organized$genres)){
  if("Drama" %in% unlist(Movies.organized$genres[i])){
    Movies$dDrama[i] <- 1
  }
}

Movies$dComedy <- 0 #(Dummy Comedy)
for (i in 1:length(Movies.organized$genres)){
  if("Comedy" %in% unlist(Movies.organized$genres[i])){
    Movies$dComedy[i] <- 1
  }
}

Movies$dThriller <- 0 #(Dummy Thriller)
for (i in 1:length(Movies.organized$genres)){
  if("Thriller" %in% unlist(Movies.organized$genres[i])){
    Movies$dThriller[i] <- 1
  }
}

Movies$dAction <- 0 #(Dummy Action)
for (i in 1:length(Movies.organized$genres)){
  if("Action" %in% unlist(Movies.organized$genres[i])){
    Movies$dAction[i] <- 1
  }
}

Movies$dAdventure <- 0 #(Dummy Adventure)
for (i in 1:length(Movies.organized$genres)){
  if("Adventure" %in% unlist(Movies.organized$genres[i])){
    Movies$dAdventure[i] <- 1
  }
}
```

##### 2.2.2 Language
```{r Create dummy variables for languages}
Movies$dEnglish <- 0 #(Dummy en)
Movies$dEnglish <- ifelse(Movies.organized$original_language == 'en', 1,0)
```

##### 2.2.3 Company
```{r Split Company, warning = FALSE}
company_count_nums <- Count.Category(Movies.organized$production_companies)
companyNames.factor <- as.factor(head(names(sort(company_count_nums,decreasing=TRUE)),5))

split.company <- spe_category(Movies.organized$production_companies)
Movies.organized$production_companies <- split.company # replace production company with a readable form
```

```{r Create dummy variables for Companies}
Movies$dRelativity <- 0 #(Dummy Relativity Media)
for (i in 1:length(Movies.organized$production_companies)){
  if("Relativity Media" %in% unlist(Movies.organized$production_companies[i])){
    Movies$dRelativity[i] <- 1
  }
}

Movies$dUniversal <- 0 #(Dummy Universal Pictures)
for (i in 1:length(Movies.organized$production_companies)){
  if("Universal Pictures" %in% unlist(Movies.organized$production_companies[i])){
    Movies$dUniversal[i] <- 1
  }
}
Movies$dWarner <- 0 #(Dummy Warner Bros.)
for (i in 1:length(Movies.organized$production_companies)){
  if("Warner Bros." %in% unlist(Movies.organized$production_companies[i])){
    Movies$dWarner[i] <- 1
  }
}

Movies$dColumbia <- 0 #(Dummy Columbia Pictures)
for (i in 1:length(Movies.organized$production_companies)){
  if("Columbia Pictures" %in% unlist(Movies.organized$production_companies[i])){
    Movies$dColumbia[i] <- 1
  }
}

Movies$dDune <- 0 #(Dummy Dune Entertainment)
for (i in 1:length(Movies.organized$production_companies)){
  if("Dune Entertainment" %in% unlist(Movies.organized$production_companies[i])){
    Movies$dDune[i] <- 1
  }
}
```

# Part3: Regressions

### 3.1 Linear Regression

##### 3.1.1 Full Model & Assumptions Checking
```{r Full model, warning=FALSE}
Full <- lm(popularity~.,data=Movies)
Full.result <- summary(Full)
par(mfrow=c(2,2))
plot(Full)

p_outlier <- rstandard(Full)[rstandard(Full) < -3 | rstandard(Full) > 3] # check for outliers that are three standard deviations away
outlier_index <- names(p_outlier)[2:length(p_outlier)]
potential_outlier <- data.frame(Movies.organized[outlier_index,])


Movies.t <- cbind(Movies.organized$title,Movies)
for(i in 1:nrow(Movies.t)){ # remove outliers
  if(Movies.t$`Movies.organized$title`[i] %in% potential_outlier$title){
    Movies.t <- Movies.t[-c(i),]
  }
}

Movies <- Movies.t[,2:ncol(Movies.t)]
Full <- lm(popularity~.,data=Movies) # full model after removing outliers
Full.result <- summary(Full)
par(mfrow=c(2,2))
plot(Full)
full.adjr <- Full.result$adj.r.squared
Full.anova <- anova(Full)
Full.SSE <- Full.anova[nrow(Full.anova),2]
n = nrow(Movies)
k = 16+2
Full.AIC = n*log(Full.SSE/n)+2*k
Full.BIC = n*log(Full.SSE/n)+k*log(n)
```


##### 3.1.2 Reduced Model by T-test
```{r Reduce model using stepwise selection}
# Backward Stepwise Selection
backward <- step(Full,direction='backward',scope=formula(Full),trace=0)
back.anova <- backward$anova
# Forward Stepwise Selection
Empty <- lm(popularity~1,data=Movies)
forward <- step(Empty,direction='forward',scope=formula(Full),trace=0)
forward.anova <- forward$anova
# Both Stepwise Selection
both <- step(Empty,direction='both',scope=formula(Full),trace=0)
bi.anova <- both$anova
# Reduced Models
Reduced <- lm(formula = popularity ~ vote_count + budget + runtime + vote_average + dThriller + dDune + dEnglish + dDrama + dColumbia, data = Movies)
par(mfrow=c(2,2))
Reduced.result <- summary(Reduced)
Reduced.adjr <- Reduced.result$adj.r.squared

Reduced.anova <- anova(Reduced)
Reduced.SSE = Reduced.anova[nrow(Reduced.anova),2]
n = nrow(Movies)
k = 9+2
Reduced.AIC = n*log(Reduced.SSE/n)+2*k
Reduced.BIC = n*log(Reduced.SSE/n)+k*log(n)
```

##### 3.1.3 Distribution of Y & Transformation Decision
```{r Histogram of popularity}
hist(Movies$popularity,breaks = 100,main="Histogram of Popularity")
# The distribution of Y is right-skewed. (square root, cube root, and log)
popularity.result <- summary(Movies$popularity)
# The range is 1.729 - 150.438.
```

##### 3.1.4 Box-Cox Transformation
```{r box cox transformation, warning=FALSE}
library(MASS)
bc <- boxcox(Full)
lambda <- bc$x[which.max(bc$y)] # find lambda

Movies.boxcox <- Movies # create new data set for box cox transformation
Movies.boxcox$popularity <- (Movies$popularity^lambda-1)/lambda
popularity_boxcox <- (Movies$popularity^lambda-1)/lambda

boxcox.Full <- lm(popularity~.,data=Movies.boxcox) # regression on full model 

Full.boxcox.result <- summary(boxcox.Full)
Full.boxcox.adjr <- Full.boxcox.result$adj.r.squared
par(mfrow=c(2,2))
plot(boxcox.Full)

boxcox.full.anova <- anova(boxcox.Full)
Full.SSE.boxcox = boxcox.full.anova[nrow(boxcox.full.anova),2]
n = nrow(Movies)
k = 16+2
Full.AIC.boxcox = n*log(Full.SSE.boxcox/n)+2*k
Full.BIC.boxcox = n*log(Full.SSE.boxcox/n)+k*log(n)

Empty.boxcox <- lm(popularity~1,data=Movies.boxcox)
forward.boxcox <- step(Empty.boxcox,direction='forward',
                    scope=formula(boxcox.Full),trace=0)
forward.boxcox.anova <- forward.boxcox$anova

boxcox.Reduced <- lm(formula = popularity ~ vote_count + budget + runtime + vote_average + dThriller + dEnglish + dDrama + dDune + dColumbia, data = Movies.boxcox) #regression

boxcox.Reduced.result <- summary(boxcox.Reduced)
Reduced.boxcox.adjr <- boxcox.Reduced.result$adj.r.squared
par(mfrow=c(2,2))
plot(boxcox.Reduced)

boxcox.reduce.anova <- anova(boxcox.Reduced)
Reduced.SSE.boxcox = boxcox.reduce.anova[nrow(boxcox.reduce.anova),2]
n = nrow(Movies)
k = 9+2
Reduced.AIC.boxcox = n*log(Reduced.SSE.boxcox/n)+2*k
Reduced.BIC.boxcox = n*log(Reduced.SSE.boxcox/n)+k*log(n)
```

##### 3.1.5 Log Transformation
```{r log transformation, warning=FALSE}
# due to unequal variance, transformation of y is used.
Movies.log <- Movies
Movies.log$popularity <- log(Movies$popularity)
# full model with log
Full.log <- lm(popularity~.,data=Movies.log)
Full.log.result <- summary(Full.log)
Full.log.adjr <- Full.log.result$adj.r.squared
par(mfrow=c(2,2))
plot(Full.log)

log.full.anova <- anova(Full.log)
Full.SSE.log = log.full.anova[nrow(log.full.anova),2]
n = nrow(Movies)
k = 16+2
Full.AIC.log = n*log(Full.SSE.log/n)+2*k
Full.BIC.log = n*log(Full.SSE.log/n)+k*log(n)

# reduced model with log
Empty.log <- lm(popularity~1,data=Movies.log)
forward.log <- step(Empty.log,direction='forward',scope=formula(Full.log),trace=0)
forward.log.anova <- forward.log$anova

Reduced.log <- lm(formula = popularity ~ vote_count + dEnglish + budget + revenue + dThriller + runtime + vote_average + dDrama + dColumbia + dComedy, data = Movies.log) # regression

Reduced.log.result <- summary(Reduced.log)
Reduced.log.adjr <- Reduced.log.result$adj.r.squared

log.reduce.anova <- anova(Reduced.log)
Reduced.SSE.log = log.reduce.anova[nrow(log.reduce.anova),2]
n = nrow(Movies)
k = 10+2
Reduced.AIC.log = n*log(Reduced.SSE.log/n)+2*k
Reduced.BIC.log = n*log(Reduced.SSE.log/n)+k*log(n)
```

##### 3.1.6 Square Root Transformation
```{r square root transformation, warning=F}
Movies.sqrt <- Movies
Movies.sqrt$popularity <- sqrt(Movies.sqrt$popularity)
# full model with square root
Full.sqrt <- lm(popularity~.,data=Movies.sqrt) # regression
Full.sqrt.result <- summary(Full.sqrt)
Full.sqrt.adjr <- Full.sqrt.result$adj.r.squared
par(mfrow=c(2,2))
plot(Full.sqrt)

sqrt.full.anova <- anova(Full.sqrt)
Full.SSE.sqrt = sqrt.full.anova[nrow(sqrt.full.anova),2]
n = nrow(Movies)
k = 16+2
Full.AIC.sqrt = n*log(Full.SSE.sqrt/n)+2*k
Full.BIC.sqrt = n*log(Full.SSE.sqrt/n)+k*log(n)

# reduced model with square root
Empty.sqrt <- lm(popularity~1,data=Movies.sqrt)
forward.sqrt <- step(Empty.sqrt,direction='forward',scope=formula(Full.sqrt),trace=0)
forward.sqrt.anova <- forward.sqrt$anova

Reduced.sqrt <- lm(formula = popularity ~ vote_count + budget + revenue + runtime + dComedy + vote_average + dEnglish + dDrama + dColumbia + dDune + dThriller, data = Movies.sqrt) #regression

Reduced.sqrt.result <- summary(Reduced.sqrt)
Reduced.sqrt.adjr <- Reduced.sqrt.result$adj.r.squared
par(mfrow=c(2,2))
plot(Reduced.sqrt)

sqrt.reduce.anova <- anova(Reduced.sqrt)
Reduced.SSE.sqrt = sqrt.reduce.anova[nrow(sqrt.reduce.anova),2]
n = nrow(Movies)
k = 11+2
Reduced.AIC.sqrt = n*log(Reduced.SSE.sqrt/n)+2*k
Reduced.BIC.sqrt = n*log(Reduced.SSE.sqrt/n)+k*log(n)
```

##### 3.1.7 Decision of Transformation (combined with 3.1.3)
```{r decision of transformation, warning= FALSE}
library(rcompanion)
par(mfrow = c(1,3))
plotNormalHistogram(popularity_boxcox,main='Box-Cox Transformation')
plotNormalHistogram(Movies.log$popularity,main='Log Transformation')
plotNormalHistogram(Movies.sqrt$popularity,main='Square Root Transformation')
```

##### 3.1.8 Decision of Linear Regression by AIC and BIC
```{r summary of linear regressions}
linear.adjr2 = c(full.adjr,Reduced.adjr)
boxcox.adjr2 = c(Full.boxcox.adjr,Reduced.boxcox.adjr)
log.adjr2 = c(Full.log.adjr,Reduced.log.adjr)
sqrt.adjr2 = c(Full.sqrt.adjr,Reduced.sqrt.adjr)
r2 = data.frame(cbind(linear.adjr2,boxcox.adjr2,log.adjr2,sqrt.adjr2))
rownames(r2) = c('Full Model','Reduced Model')

linear.AIC = c(Full.AIC,Reduced.AIC)
boxcox.AIC = c(Full.AIC.boxcox,Reduced.AIC.boxcox)
log.AIC = c(Full.AIC.log,Reduced.AIC.log)
sqrt.AIC = c(Full.AIC.sqrt,Reduced.AIC.sqrt)
AIC = data.frame(cbind(linear.AIC,boxcox.AIC,log.AIC,sqrt.AIC))
rownames(AIC) = c('Full Model','Reduced Model')

linear.BIC = c(Full.BIC,Reduced.BIC)
boxcox.BIC = c(Full.BIC.boxcox,Reduced.BIC.boxcox)
log.BIC = c(Full.BIC.log,Reduced.BIC.log)
sqrt.BIC = c(Full.BIC.sqrt,Reduced.BIC.sqrt)
BIC = data.frame(cbind(linear.BIC,boxcox.BIC,log.BIC,sqrt.BIC))
rownames(BIC) = c('Full Model','Reduced Model')
```

### 3.2 LASSO Regression
```{r LASSO regression, warning=FALSE}
library(glmnet)
Movies.Lasso <- subset(Movies, select = -c(popularity)) 
y <- Movies$popularity
x <- data.matrix(Movies.Lasso)

Movies.lasso <- cv.glmnet(x,y,alpha=1,standardize=TRUE)
lambda.best.l <- Movies.lasso$lambda.min
lasso.model.best <- glmnet(x,y,alpha=1,lambda=lambda.best.l,standardize=TRUE)
lasso.models <- glmnet(x,y,alpha=1)

par(mfrow = c(1,2))
plot(Movies.lasso)
plot(lasso.models,xvar="lambda")
par(mfrow = c(1,1))
```

### 3.3 Ridge Regression
```{r Ridge regression}
Movies.ridge <- cv.glmnet(x,y,alpha=0)
lambda.best.r <- Movies.ridge$lambda.min
ridge.model.best <- glmnet(x,y,alpha=0,lambda=lambda.best.r)
ridge.models <- glmnet(x,y,alpha=0)
coef(ridge.model.best)
par(mfrow = c(1,2))
plot(Movies.ridge)
plot(ridge.models,xvar="lambda")
par(mfrow = c(1,1))
```

# Part4: Model Selection

### 4.1 R-squared, MSE, AIC, and BIC
```{r model selection, warning=FALSE}
#Lasso
predicted.l <- predict(lasso.model.best,s=lambda.best.l,newx=x)
k <- lasso.model.best$df
n <- lasso.model.best$nobs
SSE <- sum((predicted.l - y)^2)
MSE <- SSE/(n-k)
SSTO <- sum((y - mean(y))^2)
R_squared <- 1-SSE/SSTO
Lasso.result <- data.frame(MSE=MSE,SSE=SSE,R_squared=R_squared)
#Ridge
predicted.r <- predict(ridge.model.best,s=lambda.best.r,newx=x)
k <- ridge.model.best$df
n <- ridge.model.best$nobs
SSE <- sum((predicted.r - y)^2)
MSE <- SSE/(n-k)
SSTO <- sum((y - mean(y))^2)
R_squared <- 1-SSE/SSTO
Ridge.result <- data.frame(MSE=MSE,SSE=SSE,R_squared=R_squared)
#Reduced Log Regression
Reduced.SSE.log = log.reduce.anova[nrow(log.reduce.anova),2]
Reduced.MSE.log = log.reduce.anova[nrow(log.reduce.anova),3]
# summary(Reduced)
Reduced.R2.log <- Reduced.log.result$r.squared
k <- 10+2
n <- 148
Log.result <- data.frame(MSE=Reduced.SSE.log,SSE=Reduced.MSE.log,R_squared=Reduced.R2.log)

```

### 4.2 Model Selection Conclusion
```{r}
selection_table <- rbind(Lasso.result=Lasso.result,
                         Ridge.result=Ridge.result,
                         Log.result=Log.result)
```

# Part5: Movie Recommendation Function
```{r}
Movies_recommendation <- function(){
  print(c('Here is the list of genres:',names(genre_count_nums)))
  Genre = readline(prompt="Enter Preferred Genre: ")
  recommend.index = c()
  for (i in 1:nrow(Movies.organized)){
    if (Genre %in% unlist(Movies.organized$genres[i])){
      recommend.index = c(recommend.index,i)
    }
  }
  Movies.recommend = data.frame(Movies.organized[recommend.index,])
  Movies.recommend = Movies.recommend[order(-Movies.recommend$popularity),]
  Movie = head(Movies.recommend$title,3)
  print(c("The most popular three movies are:",Movie))
}
```